### 2025-05-11 conteudo/background/linux.tex
add sowrk 11/05

One of the subsystems within the Linux kernel is the AMD Display Driver, which
is responsible for implementing the drivers needed for AMD GPUs to function
correctly in a Linux environment. The maintainers of this subsystem have
reported a practical issue: a significant amount of duplicated code, which
complicates the maintenance of this subsystem. This problem serves as the
starting point for our research, with the AMD Display Driver as our primary
object of study as we aim to understand and mitigate code duplication in the
Linux kernel. The AMD Display Driver subsystem contains more then 391000 lines 
of code and more then 1089 code files
\footnote{ 
Data extracted directly from the Linux repository with the help of the cloc tool.
\citep{cloc}
}, 
making it a considerable component within the Linux kernel. Like other 
subsystems, discussions about the AMD Display Driver are conducted on its 
designated mailing list, which can be accessed here: 
\url{https://lists.freedesktop.org/mailman/listinfo/amd-gfx}.

\todo[inline]{Require an introduction about IIO drivers.}

>>>>

One of the subsystems within the Linux kernel is the AMD Display driver, which
is responsible for implementing the drivers needed for AMD GPUs to function
correctly in a Linux environment. The maintainers of this subsystem have
reported a practical issue: a significant amount of duplicated code, which
complicates the maintenance of this subsystem. This problem serves as the
starting point for our research, with the AMD Display Driver as our primary
object of study as we aim to understand and mitigate code duplication in the
Linux kernel. The AMD Display driver subsystem contains more then 391000 lines 
of code and more then 1089 code files
\footnote{ 
Data extracted directly from the Linux repository with the help of the cloc tool.
\citep{cloc}
}, 
making it a considerable component within the Linux kernel. Like other 
subsystems, discussions about the AMD Display driver are conducted on its 
designated mailing list, which can be accessed here: 
\url{https://lists.freedesktop.org/mailman/listinfo/amd-gfx}.

Another subsytem within the Linux kernel explored in this work is the Industrial
I/O subsytem (IIO). From the Linux kernel documentation, ''The main purpose of the 
Industrial I/O subsystem (IIO) is to provide support for devices that in some sense 
perform either analog-to-digital conversion (ADC) or digital-to-analog conversion (DAC) or both``\citep{iiodoc}.
The IIO subsystem contains more then X lines of code and more then Y code files 
\footnote{
Data extracted directly from the Linux repository with the help of the clol tool. \citep{cloc}
}.
We opted to include the IIO subsystem in our experiments to ensure our findings are not limited to a single subsystem. 
Additionally, we have an established point of contact within the IIO subsystem to facilitate this work.
. Like other subsystems, discussions about the IIO subsytem are conducted on its 
designated mailing list, which can be found here: 
\url{https://subspace.kernel.org/vger.kernel.org.html}.

### 2025-05-11 conteudo/introduction/intro.tex
add sowrk 11/05

\todo[inline]{Maybe talk about the tests on IIO software, as the test is not tested only on AMD Display driver anymore. Maybe mention the tool explicitly.}

@@

\todo[inline]{THIS SECTION NEEDS TO BE REMADE FROM ZERO.}

Our studies and observations reveal that the code duplication problem described
in the previous section cannot be fully addressed by existing tools in
either the formal or grey literature. To fill this gap, we propose an approach
to identify and mitigate code duplications specifically within the Linux
kernel. Our approach includes the development of a tool capable of detecting
duplicated functions within the kernel and establishing strategies for
addressing these duplications.

\todo[inline]{Focus more on the tool here.}

@@

This research aims to identify patterns of software engineering techniques, 
such as refactoring methods commonly used to address code duplications. 
These patterns will serve as a guide for future contributors in mitigating 
code duplication issues within this context. To investigate our approach, 
we are developing a tool to identify duplicate code in the Linux kernel, 
explicitly targeting parts of the AMD Display driver code and the IIO driver. We are then refactoring 
portions of this code and submitting patches to maintainers, providing evidence 
of the tool's accuracy in detecting duplicate functions. Our research design follows 
a multi-method approach, structured into two phases, each employing different research methods.

@@

\section**{Manuscript Structure}

This manuscript consists of five more chapters. Chapter \ref{cha:back} presents
the literature overview for code duplication detection (main definitions,
current approaches in the literature), a brief description of the Linux kernel,
and a review of refactoring methods used throughout this research. Chapter
\ref{cha:tool} presents Arkanjo, our proposed tool to detect code duplication in the
Linux kernel, describing all the main components.  Chapter \ref{cha:method}
describes the research methods selected to guide our work. Chapter
\ref{cha:results} shows the first results obtained from the research methods to evaluate
our tool and presents the current research stage and the work plan,
with the addition of a discussion about our preliminary results.

>>>>

@@

Our studies and observations reveal that the code duplication problem described
in the previous section cannot be fully addressed by existing tools in
either the formal or grey literature. To fill this gap, we propose an approach
to identify and mitigate code duplications specifically within the Linux
kernel. Our approach includes the development of a tool capable of detecting
duplicated functions and establishing strategies for
addressing these duplications. We tested the approach in the context of two
kernel components, the AMD Display Driver the the Industrial IO subsystem.

@@

This research aims to identify patterns of software engineering techniques, 
such as refactoring methods commonly used to address code duplications. 
These patterns will serve as a guide for future contributors in mitigating 
code duplication issues within this context. To investigate our approach, 
we are developing a tool to identify duplicate code. We are then refactoring 
portions of this code and submitting patches to maintainers, providing evidence 
of the tool's accuracy in detecting duplicate functions. Our research design follows 
a multi-method approach, structured into two phases, each employing different research methods.

@@

\section**{Thesis Structure}

This manuscript consists of five more chapters. Chapter \ref{cha:back} presents
the literature overview for code duplication detection (main definitions,
current approaches in the literature), a brief description of the Linux kernel and the components
explored in this work, 
and a review of refactoring methods used throughout this research. Chapter
\ref{cha:tool} presents ArKanjo, our proposed tool to detect code duplication, 
describing all the main components.  Chapter \ref{cha:method}
describes the research methods selected to guide our work. Chapter
\ref{cha:results} shows the results we had through our work, from the research methods to 
evaluate our tool and the ethnographic studies. Chapter \ref{cha:conclusion}udes this research.

### 2025-05-11 conteudo/results/participant.tex
reorginaze results files
add sowrk 11/05

\section{Final Remarks}

We have completed Phase 1 of our research, during which we developed our proposed tool and evaluated it using the BigCloneBench dataset and through empirical analyses. The results indicate our tool can identify code clone duplications within the Linux kernel.

@@

\input{conteudo/preliminar/table_patch}

@@

\input{conteudo/preliminar/cronograma}

>>>>

\section{Participant Observation Experiment}

We have completed Phase 1 of ourdd research, during which we developed our proposed tool and evaluated it using the BigCloneBench dataset and through empirical analyses. The results indicate our tool can identify code clone duplications within the Linux kernel.

@@

\input{conteudo/results/table_patch}

@@

\input{conteudo/results/cronograma}

### 2025-05-11 conteudo/results/results.tex
add sowrk 11/05

\todo[inline]{As it is primarly results, everything should be remade, the structure of the files for this section is bad.}

\todo[inline]{Remember the typos on the tables. }

\todo[inline]{Idea: The research methods should be 3 phases, one of the tool creation/testing, one for the my experiences submiting a path, and the last for the experiences of the students on the course. Maybe split in four to split the tool creation for the tool testing.}

To finalize this report, we present the current state of our research. As of the 
time of writing, we have finished \textbf{Phase I} of this research, 
which includes the implementation of 
the proposed tool and the evaluation of the tool through both a method from the 
literature and an empirical analyses. This chapter begins with the presentation of 
the results from the tool evaluation. Next, we present a preliminary experiment 
we are currently conducting to assess the viability of our Ethnographic Study. 
Finally, we outline the next steps in this research.

@@

>>>>

In this chapter we present and analyze the results found in the experiments and 
methods done through our research. Following the cronology of the research, we 
start by presenting the ArKanjo tool's evaluation through both a method from the literature
and an empirical analyses. Then, we present and discuss the experience of the initial participant
observation methodolody, where we mitigate a code duplication on the AMD Display driver and sent it as a patch.
Finally, we show and discuss our discoveries on the experiment done with students at the course in University of 
São Paulo, where we propose to students to mitigate duplications found by ArKanjo tool at IIO subsystem and the
AMD Display driver and document their founds, opinions and experiences.

@@

\input{conteudo/results/students}

### 2025-05-11 conteudo/tool/tool.tex
remove setup and change arc diagram
add sowrk 11/05

The tool operates in two main parts, the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} is responsible to perform the heavy computations to find the code duplications 
accross a codebase and produce artifact to consulte information related to the duplications in a structured form.
The \textbf{Query Responder} consumes the artifacts produced by the Preprocessor to execute the tool functionalites as requested.
This choice enable the tool to concentrate the heavy and slows parts that can be execute only once per codebase, 
thus allowing a fast performance for multiple queries related to a single codebase.
More detailed descriptions of the Setup and Query Responder processes can be found in
Sections \ref{subsec:architecture} and \ref{subsec:func}.

@@

\todo[inline]{I need to rename the Setup as Preprocessor step.}

\todo[inline]{Update the Function parser description of the C parser and Java parser}

\todo[inline]{Add the other similarity method added on the tool using git for the Code Duplication Finder.}

\todo[inline]{Update diagram and functionalites pictures (do not need sudo anymore!).}

>>>>

The tool operates in two main parts, the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} is responsible to perform the heavy computations to find the code duplications 
accross a codebase and produce artifact to consulte information related to the duplications in a structured form.
The \textbf{Query Responder} consumes the artifacts produced by the Preprocessor to execute the tool functionalites as requested.
This choice enable the tool to concentrate the heavy and slows parts that can be execute only once per codebase, 
thus allowing a fast performance for multiple queries related to a single codebase.
More detailed descriptions of the Preprocessor and Query Responder can be found in
Sections \ref{subsec:architecture} and \ref{subsec:func}.

@@
